---
title: "Vanilla SV"
author: "Catherine Forbes"
date: "FE2 2019"
output: html_document
---

```{r setup, include=TRUE, echo = FALSE, eval=TRUE, warning = FALSE, message = FALSE, error = FALSE, fig.height=4, fig.width=10}
knitr::opts_chunk$set(echo = FALSE, eval=TRUE, warning = FALSE, message = FALSE, error = FALSE, fig.height=4, fig.width=10, comment = NA)

library(tidyverse)
library(gridExtra)
library(moments)
library(broom)
library(kableExtra)
```

### Model specification

\begin{eqnarray}
r_{}& = & \mu +\sigma_t \varepsilon_t; \ \ \varepsilon_t  \overset{i.i.d}{\sim} N(0,1) \\
& &   \\
\ln (\sigma_t^2) & = & \alpha_0 + \alpha_1 \ln (\sigma_{t-1}^2) + v_t; \ \ v_t \overset{i.i.d}{\sim} N(0,\sigma_v^2) \\
& & \\
corr(\varepsilon_t, v_t) & = & \rho \label{svmodel} 
\end{eqnarray}

\vspace{.3cm}


## Zero correlation

Settings:

```{r echo=TRUE}
mu <- 0.0009
a0 <- -0.7824
a1 <- 0.9
sigv <- 0.405
rho <- 0.0
```

```{r}

# generate

N <- 2000
n <- 1000
set.seed(20199876)

v <- rnorm(N,0,1)
eps <- rho*v + sqrt(1-rho^2)*rnorm(N,0,1)

lnsig2 <- numeric()
rt <- numeric()

lnsig2[1] <- a0/(1-a1)
rt[1] <- mu + exp(lnsig2[1]/2)*eps[1]

for(i in 2:N){
  lnsig2[i] <- a0 + a1*lnsig2[i-1] + sigv*v[i]
  rt[i] <- mu + exp(lnsig2[i]/2)*eps[i]
}

# plot

df <- tibble(day=1:n, logsig2=lnsig2[(N-n+1):N], rt=rt[(N-n+1):N])
df <- df %>% mutate(sig = exp(logsig2/2), rt2 = rt^2)

p1 <- df %>% ggplot(aes(x=day,y=rt)) + geom_line(col="blue") + ggtitle("returns") + ylab(expression(r[t]))
p2 <- df %>% ggplot(aes(x=day,y=logsig2)) + geom_line(col="blue") + ggtitle("log variance") + ylab(expression(ln(sigma[t]^2)))
p3 <- df %>% ggplot(aes(x=day,y=sig)) + geom_line(col="blue") + ggtitle("volatility") + ylab(expression(sigma[t]))
p4 <- df %>% ggplot(aes(x=rt, y=..density..)) + geom_histogram(col="blue",bins=50) + ggtitle("histogram of returns") + ylab(expression(r[t]))

grid.arrange(p1, p4, p2, p3, nrow=2)

```

### Descriptive stats

```{r}

jb <- function(x){
  n <- length(x)
  sighat <- sqrt((n-1)/n)*sd(x)
  eps <- (x - mean(x))/sighat
  sk <- mean(eps^3)
  kt <- mean(eps^4)
  ekt <- kt - 3
  jb <- (n/6)*(sk^2 + (ekt^2)/4)
  return(jb)
  }

tbl <- df %>% summarise(min=min(rt), q25=quantile(rt,0.25),med=quantile(rt,0.5),q75=quantile(rt,0.75), max=max(rt), mean=mean(rt), sd=sd(rt), sk=skewness(rt), kt=kurtosis(rt), JB=jb(rt), pval=pchisq(jb(rt),df=2, lower.tail=FALSE) ) 

tbl %>% kable() %>% kable_styling()
```


## Simultion 2: Negative correlation

Settings: 

```{r echo=TRUE}
mu <- 0.0009
a0 <- -0.7824
a1 <- 0.9
sigv <- 0.405
rho <- -0.1
```

```{r}

# generate 

N <- 5000
n <- 1000
set.seed(20199876) # use same seed

v <- rnorm(N,0,1)
eps <- rho*v + sqrt(1-rho^2)*rnorm(N,0,1)
 
lnsig2 <- numeric()
rt <- numeric()

lnsig2[1] <- a0/(1-a1)
rt[1] <- mu + exp(lnsig2[1]/2)*eps[1]

for(i in 2:N){
  lnsig2[i] <- a0 + a1*lnsig2[i-1] + sigv*v[i]
  rt[i] <- mu + exp(lnsig2[i]/2)*eps[i]
}

# plot

df <- tibble(day=1:n, logsig2=lnsig2[(N-n+1):N], rt=rt[(N-n+1):N])
df <- df %>% mutate(sig = exp(logsig2/2), rt2 = rt^2)

p1 <- df %>% ggplot(aes(x=day,y=rt)) + geom_line(col="blue") + ggtitle("returns") + ylab(expression(r[t]))
p2 <- df %>% ggplot(aes(x=day,y=logsig2)) + geom_line(col="blue") + ggtitle("log variance") + ylab(expression(ln(sigma[t]^2)))
p3 <- df %>% ggplot(aes(x=day,y=sig)) + geom_line(col="blue") + ggtitle("volatility") + ylab(expression(sigma[t]))
p4 <- df %>% ggplot(aes(x=rt, y=..density..)) + geom_histogram(col="blue",bins=50) + ggtitle("histogram of returns") + ylab(expression(r[t]))

grid.arrange(p1, p4, p2, p3, nrow=2)

```

### Descriptive stats

```{r}
tbl <- df %>% summarise(min=min(rt), q25=quantile(rt,0.25),med=quantile(rt,0.5),q75=quantile(rt,0.75), max=max(rt), mean=mean(rt), sd=sd(rt), sk=skewness(rt), kt=kurtosis(rt), JB=jb(rt), pval=pchisq(jb(rt),df=2, lower.tail=FALSE) ) 
 
tbl %>% kable() %>% kable_styling()
```




## Simulation 3: Positive correlation

Settings:

```{r echo=TRUE}
mu <- 0.0009
a0 <- -0.7824
a1 <- 0.9
sigv <- 0.405
rho <- 0.1
```

```{r}

# generate 

N <- 5000
n <- 1000
set.seed(20199876) # use same seed

v <- rnorm(N,0,1)
eps <- rho*v + sqrt(1-rho^2)*rnorm(N,0,1)


lnsig2 <- numeric()
rt <- numeric()

lnsig2[1] <- a0/(1-a1)
rt[1] <- mu + exp(lnsig2[1]/2)*eps[1]

for(i in 2:N){
  lnsig2[i] <- a0 + a1*lnsig2[i-1] + sigv*v[i]
  rt[i] <- mu + exp(lnsig2[i]/2)*eps[i]
}

# plot

df <- tibble(day=1:n, logsig2=lnsig2[(N-n+1):N], rt=rt[(N-n+1):N])
df <- df %>% mutate(sig = exp(logsig2/2), rt2 = rt^2)

p1 <- df %>% ggplot(aes(x=day,y=rt)) + geom_line(col="blue") + ggtitle("returns") + ylab(expression(r[t]))
p2 <- df %>% ggplot(aes(x=day,y=logsig2)) + geom_line(col="blue") + ggtitle("log variance") + ylab(expression(ln(sigma[t]^2)))
p3 <- df %>% ggplot(aes(x=day,y=sig)) + geom_line(col="blue") + ggtitle("volatility") + ylab(expression(sigma[t]))
p4 <- df %>% ggplot(aes(x=rt, y=..density..)) + geom_histogram(col="blue",bins=50) + ggtitle("histogram of returns") + ylab(expression(r[t]))

grid.arrange(p1, p4, p2, p3, nrow=2)

```

### Descriptive statistics for the returns

```{r}
tbl <- df %>% summarise(min=min(rt), q25=quantile(rt,0.25),med=quantile(rt,0.5),q75=quantile(rt,0.75), max=max(rt), mean=mean(rt), sd=sd(rt), sk=skewness(rt), kt=kurtosis(rt), JB=jb(rt), pval=pchisq(jb(rt),df=2, lower.tail=FALSE) ) 

tbl %>% kable() %>% kable_styling()
```


## Simulation 4: Greater volatility

Settings:

```{r echo=TRUE}
mu <- 0.0009
a0 <- -0.7824
a1 <- 0.9
sigv <- 0.6
rho <- 0.1
```

```{r}

# generate 

N <- 5000
n <- 1000
set.seed(20199876)

v <- rnorm(N,0,1)
eps <- rho*v + sqrt(1-rho^2)*rnorm(N,0,1)


lnsig2 <- numeric()
rt <- numeric()

lnsig2[1] <- a0/(1-a1)
rt[1] <- mu + exp(lnsig2[1]/2)*eps[1]

for(i in 2:N){
  lnsig2[i] <- a0 + a1*lnsig2[i-1] + sigv*v[i]
  rt[i] <- mu + exp(lnsig2[i]/2)*eps[i]
}

# plot

df <- tibble(day=1:n, logsig2=lnsig2[(N-n+1):N], rt=rt[(N-n+1):N])
df <- df %>% mutate(sig = exp(logsig2/2), rt2 = rt^2)

p1 <- df %>% ggplot(aes(x=day,y=rt)) + geom_line(col="blue") + ggtitle("returns") + ylab(expression(r[t]))
p2 <- df %>% ggplot(aes(x=day,y=logsig2)) + geom_line(col="blue") + ggtitle("log variance") + ylab(expression(ln(sigma[t]^2)))
p3 <- df %>% ggplot(aes(x=day,y=sig)) + geom_line(col="blue") + ggtitle("volatility") + ylab(expression(sigma[t]))
p4 <- df %>% ggplot(aes(x=rt, y=..density..)) + geom_histogram(col="blue",bins=50) + ggtitle("histogram of returns") + ylab(expression(r[t]))


grid.arrange(p1, p4, p2, p3, nrow=2)

```

### Descriptive statistics for the returns

```{r}
tbl <- df %>% summarise(min=min(rt), q25=quantile(rt,0.25),med=quantile(rt,0.5),q75=quantile(rt,0.75), max=max(rt), mean=mean(rt), sd=sd(rt), sk=skewness(rt), kt=kurtosis(rt), JB=jb(rt), pval=pchisq(jb(rt),df=2, lower.tail=FALSE) ) 

tbl %>% kable() %>% kable_styling()
```

\vspace{.5cm}


